\Huge \bf \sc Calibration of an inertial measurement unit

### 2013-07-29 latex/paper_sensores/paper.tex
paper terminado y alguna cosa del de los sensores

@@

\section{Introduction}
The development of the Micro-electromechanical Systems(MEMS) technology has allowed to manufacture many low-cost chip-sensors, such as accelerometers, gyroscopes and magnetometers. Those chips have been adopted in many applications, for instance Inertial Navigation Systems (INS)\cite{bib:paper}. However this sensors have many error sources, thus they must be calibrated before being used and they should be re-calibrated periodically for any precision application. The different sources of error will be analysed in more detail in section \refp{sec:modelo} and the implications of those errors will be derived.\\

The calibration proposed in this paper is tested on a 3-axis accelerometer ADXL345 of Analog Devices and a 3-axis gyroscope ITG-3200  of InvenSense, yet the model developed and the methodology can be used in other devices because the model is based on common characteristics of MEMS sensors. The sensors named before are included in a Inertial Measurement Unit(IMU): The Mongoose (figure \refp{fig:mongoose}). \\

@@

As it was stated before, there exist many error sources in MEMS devices. In \cite{bib:calib_imu} two of this sources are mentioned: the nonlinear response of the sensors and the non-orthogonality of the axis of the sensor. In addition we can observe also that it exist an electric noise in the measures (figure \ref{fig:noise}) and a dependence of the temperature.\\
\vspace{-10pt}
\begin{figure}[h!]
  \centering
  \subfloat[Accelerometer data $\hat{k}$]{\label{fig:angulos_1}
  		\includegraphics[width=1\columnwidth]
  			{./pics_paper/acc_noise.pdf}}\\[-5pt]
  			\vspace{-5pt}
  \subfloat[Gyroscope data $\hat{j}$]{\label{fig:angulos_2} 
  		\includegraphics[width=1\columnwidth]
  			{./pics_paper/gyro_noise.pdf}}
  \caption{\textbf{Noise -} Measures from accelerometer and gyroscope in static equilibrium.}
  \label{fig:noise}
\end{figure}

@@

In figure \refp{fig:noise} we can observe that the electrical noise of the devices is typically 2 bits, this error (according to the datasheets \cite{bib:acc_data} and \cite{bib:gyro_data}) correspond to $7.8 mg$ and $0.14 ^{\circ}/s $. The error is not biased, thus we not need to consider it, because several samples are used in any real application and the effect vanishes.\\

@@

\subsection{Accelerometer}
Due to construction issues the sensitivity axis of the device are generally not orthogonal. Lets define a platform system as shown in figure \refp{fig:mongoose}. Let $\mathbf{a}^a$ be the true acceleration (meaning the theoretical acceleration measured in $ms^{-2}$) expressed in the sensitivity axis of the accelerometer and let $\mathbf{a}^p$ be the theoretical true acceleration (also measured in $ms^{-2}$) expressed in the platform system. We model this fact by the following linear transformation between the two acceleration vectors:

\begin{equation}
\mathbf{a}^p = \mathbf{T}_a^p\mathbf{a}^a, \; \mathbf{T}_a^p = \left(\begin{array}{ccc}
1 & - \alpha_{yz} & \alpha_{zy}\\
\alpha_{xz} & 1 & - \alpha_{zx}\\
-\alpha_{xy} & \alpha_{yx} & 1
\end{array}\right)
\label{eq:nonorto}
\end{equation}

In equation \refp{eq:nonorto}, scalars $\alpha_{ij}$ represents the rotation of the $i-th$ sensitivity axis of the accelerometer over the $j-th$ axis of the platform system. In figure \refp{fig:nonorto} this relation can be observed graphically. Since those errors are due to the manufacturing process we will assume that they will remain constant long enough.

@@

As it was stated before, we are considering a linear model between the acceleration measured in the sensitivity axis and the true acceleration in the same system. Thus we have:

\begin{equation}
\tilde{\mathbf{a}}^a = \mathbf{K}_a\mathbf{a}^a + \mathbf{b}_a
\label{eq:modlineal}
\end{equation}

In equation \refp{eq:modlineal}, $\tilde{\mathbf{a}}^a$ is the measure obtained by the accelerometer (measured in bits), $\mathbf{K}_a$ is a diagonal matrix that represents the sensitivity of each axis and $\mathbf{b}_a$ is a vector that represent the bias of each axis. Those parameters may variate as the temperature changes. However, in this stage we are going to consider them constant. Using equations \refp{eq:nonorto} and \refp{eq:modlineal} we can establish a model of the accelerometer:

@@

\section{Calibration method proposed}
The problem of calibration is to establish the values for the unknown parameters of the model that adjust ``the better'' a certain set of data. This criterion will be defined in section \refp{sec:param}.

@@

First of all, we attached the platform to a wooden cube. This allows us to perform $90^\circ$ rotations, thus if we had an horizontal surface we can take measures with the gravity vector aligned to any of three axis. The horizontal surface was assured using a three leg adjustable table (see figure \refp{fig:mesa}). In addition this table allows us to perform rotations of known angles using a scaled semicircle and a pendulum as it is shown also in figure \refp{fig:mesa}. The rotations over the second axis were made using a carpenter's square, therefore we have a proper $45^\circ$ rotation.\\

@@

In this case the constant angular speed was provided with a turntable ($33 rev/min$). Its angular speed was measured just to be sure that it was working properly. Different orientations were obtained using the wooden cube mentioned in section \refp{subsec:acc},to assure $90^\circ$ rotations and pairs of carpenter's squares to assure $30^\circ$, $45^\circ$ and $60^\circ$ rotations. In figure \refp{fig:setup_gyro} is shown the setup that was mounted over the turntable.\\

@@

To evaluate the behavior of the measurement of the accelerometer and compensate the error caused by temperature, the experiment described in Section \refp{sec:param_temp} is carried out. It can be divided in two separate parts, the first is heating up the sensor using a hair dryer up to $48^oC$ and the second part is letting the sensor cool down to normal operating temperature. The two separated parts of the experiment are shown in figure \refp{fig:temp_res}.

@@

In figure \refp{fig:resultado_temp} is shown the behavior of the calibrated measurement of the accelerometer in z-axis while the sensor is heated up, both with and without temperature compensation. It is clearly shown that the compensated temperature is much more reliable than the other one, that is varying a 3\% of the highest value. Otherwise, in figure \refp{fig:bajada} the second part of the experiment is shown, letting the sensor cool down up to normal operating temperature. Two different magnitudes are shown at the same time: the three accelerations and the temperature, thus two axes are needed to be shown: on left side the acceleration in $ms^{-2}$, and on right side the temperature in $^oC$.\\
In figure \refp{fig:resultado_temp} can be observed that at about the end of the experiment the sensor output is $9.25ms^{-2}$ instead of $9.81ms^{-2}$, which means an error of $0.56ms^{-2}$. Such an error in acceleration measurements can cause several errors while trying to control the quadrotor, thus is vital that temperature compensation is performed.

>>>>

\usepackage{balance}

@@

\section{Introduction}
The development of the Micro-electromechanical Systems (MEMS) technology has allowed to manufacture many low-cost chip-sensors, such as accelerometers, gyroscopes and magnetometers. Those chips have been adopted in many applications, for instance Inertial Navigation Systems (INS), \cite{bib:paper}. However this sensors have many error sources, thus, they must be calibrated before being used and they should be re-calibrated periodically for any precision application. The different sources of error will be analysed in more detail in Section \refp{sec:modelo} and the implications of those errors will be derived.\\

The calibration proposed in this paper is tested on a 3-axis accelerometer ADXL345 of Analog Devices and a 3-axis gyroscope ITG-3200  of InvenSense, yet the model developed and the methodology can be used in other devices because the model is based on common characteristics of MEMS sensors. The sensors named before are included in a Inertial Measurement Unit(IMU): The Mongoose (Figure \refp{fig:mongoose}). \\

@@

As it was stated before, there exist many error sources in MEMS devices. In \cite{bib:calib_imu} two of this sources are mentioned: the nonlinear response of the sensors and the non-orthogonality of the axis of the sensor. In addition we can observe also that it exist an electric noise in the measures (Figure \ref{fig:noise}) and a dependence of the temperature.\\
\vspace{-10pt}
\begin{figure}[h!]
  \centering
  \subfloat[Accelerometer data $\hat{k}$]{\label{fig:angulos_1}
  		\includegraphics[width=1\columnwidth]
  			{./pics_paper/acc_noise.pdf}}\\[-5pt]
  			\vspace{-5pt}
  \subfloat[Gyroscope data $\hat{j}$]{\label{fig:angulos_2} 
  		\includegraphics[width=1\columnwidth]
  			{./pics_paper/gyro_noise.pdf}}
  \caption{\textbf{Noise -} Measures from accelerometer and gyroscope in static equilibrium.}
  \label{fig:noise}
\end{figure}

@@

In Figure \refp{fig:noise} we can observe that the electrical noise of the devices is typically 2 bits, this error (according to the datasheets \cite{bib:acc_data} and \cite{bib:gyro_data}) correspond to $7.8 mg$ and $0.14 ^{\circ}/s $. The error is not biased, thus we not need to consider it, because several samples are used in any real application and the effect vanishes.\\

@@

\subsection{Accelerometer}
Due to construction issues the sensitivity axis of the device are generally not orthogonal. Lets define a platform system as shown in Figure \refp{fig:mongoose}. Let $\mathbf{a}^a$ be the true acceleration (meaning the theoretical acceleration measured in $ms^{-2}$) expressed in the sensitivity axis of the accelerometer and let $\mathbf{a}^p$ be the theoretical true acceleration (also measured in $ms^{-2}$) expressed in the platform system. We model this fact by the following linear transformation between the two acceleration vectors:
\begin{equation}
\mathbf{a}^p = \mathbf{T}_a^p\mathbf{a}^a, \; \mathbf{T}_a^p = \left(\begin{array}{ccc}
1 & - \alpha_{yz} & \alpha_{zy}\\
\alpha_{xz} & 1 & - \alpha_{zx}\\
-\alpha_{xy} & \alpha_{yx} & 1
\end{array}\right)
\label{eq:nonorto}
\end{equation}
In equation \refp{eq:nonorto}, scalars $\alpha_{ij}$ represents the rotation of the $i-th$ sensitivity axis of the accelerometer over the $j-th$ axis of the platform system. In Figure \refp{fig:nonorto} this relation can be observed graphically. Since those errors are due to the manufacturing process we will assume that they will remain constant long enough.

@@

As it was stated before, we are considering a linear model between the acceleration measured in the sensitivity axis and the true acceleration in the same system. Thus we have:
\begin{equation}
\tilde{\mathbf{a}}^a = \mathbf{K}_a\mathbf{a}^a + \mathbf{b}_a
\label{eq:modlineal}
\end{equation}
In equation \refp{eq:modlineal}, $\tilde{\mathbf{a}}^a$ is the measure obtained by the accelerometer (measured in bits), $\mathbf{K}_a$ is a diagonal matrix that represents the sensitivity of each axis and $\mathbf{b}_a$ is a vector that represent the bias of each axis. Those parameters may variate as the temperature changes. However, in this stage we are going to consider them constant. Using equations \refp{eq:nonorto} and \refp{eq:modlineal} we can establish a model of the accelerometer:

@@

\section{Calibration method proposed}
The problem of calibration is to establish the values for the unknown parameters of the model that adjust ``the better'' a certain set of data. This criterion will be defined in Section \refp{sec:param}.

@@

First of all, we attached the platform to a wooden cube. This allows us to perform $90^\circ$ rotations, thus if we had an horizontal surface we can take measures with the gravity vector aligned to any of three axis. The horizontal surface was assured using a three leg adjustable table (see Figure \refp{fig:mesa}). In addition this table allows us to perform rotations of known angles using a scaled semicircle and a pendulum as it is shown also in Figure \refp{fig:mesa}. The rotations over the second axis were made using a carpenter's square, therefore we have a proper $45^\circ$ rotation.\\

@@

In this case the constant angular speed was provided with a turntable ($33 rev/min$). Its angular speed was measured just to be sure that it was working properly. Different orientations were obtained using the wooden cube mentioned in Section \refp{subsec:acc},to assure $90^\circ$ rotations and pairs of carpenter's squares to assure $30^\circ$, $45^\circ$ and $60^\circ$ rotations. In Figure \refp{fig:setup_gyro} is shown the setup that was mounted over the turntable.\\

@@

To evaluate the behavior of the measurement of the accelerometer and compensate the error caused by temperature, the experiment described in Section \refp{sec:param_temp} is carried out. It can be divided in two separate parts, the first is heating up the sensor using a hair dryer up to $48^oC$ and the second part is letting the sensor cool down to normal operating temperature. The two separated parts of the experiment are shown in Figure \refp{fig:temp_res}.

@@

\balance
In Figure \refp{fig:resultado_temp} is shown the behavior of the calibrated measurement of the accelerometer in z-axis while the sensor is heated up, both with and without temperature compensation. It is clearly shown that the compensated temperature is much more reliable than the other one, that is varying a 3\% of the highest value. Otherwise, in Figure \refp{fig:bajada} the second part of the experiment is shown, letting the sensor cool down up to normal operating temperature. Two different magnitudes are shown at the same time: the three accelerations and the temperature, thus two axes are needed to be shown: on left side the acceleration in $ms^{-2}$, and on right side the temperature in $^oC$.\\
In Figure \refp{fig:resultado_temp} can be observed that at about the end of the experiment the sensor output is $9.25ms^{-2}$ instead of $9.81ms^{-2}$, which means an error of $0.56ms^{-2}$. Such an error in acceleration measurements can cause several errors while trying to control the quadrotor, thus is vital that temperature compensation is performed.
